---
title: "nonlinear_age"
author: "SL"
date: "April 17, 2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, mesage = FALSE, warning = FALSE)
library(tidyverse)
library(mgcv)
library(splines)
library(ggcorrplot)
library(corrplot)
library(modelr)
library(voxel)
library(gridExtra)

```

Load dataset from pipeline output
```{r load}
pfas <- read.csv("aamehs_data.csv") %>% 
   select(seqn, gender, age, race_ethnicity, hh_education, pfdea, pfhxs, me_pfosa_acoh, pfna, pfua, pfdoa, n_pfoa, sb_pfoa, n_pfos, sm_pfos, bmi) %>% 
  mutate(gender = factor(gender),
         race_ethnicity = factor(race_ethnicity),
         hh_education = factor(hh_education)) %>% 
  drop_na()


pfas_adolescent <-
  pfas %>%
  filter(age <= 19)


pfas_adult <-
  pfas %>%
  filter(age > 19)

```


## Data Summary
```{r, eval = FALSE}

pfas %>% 
  select(gender, age, race_ethnicity, hh_education, bmi, everything(), -seqn) %>% 
  summary(.) 

```


## Plots
```{r}
pfas_adolescent %>%
    select(bmi, pfdea:sm_pfos) %>%
    gather(pfdea:sm_pfos, key = "variable", value = "value") %>% 
  ggplot(aes(value, bmi)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~variable) +
    theme_bw()


pfas_adult %>%
    select(bmi, pfdea:sm_pfos) %>%
    gather(pfdea:sm_pfos, key = "variable", value = "value") %>% 
  ggplot(aes(value, bmi)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~variable) +
    theme_bw()

```

## Correlations

```{r}
cor_pfas_adolescent <-
  pfas %>%
    select(-bmi, -(seqn:hh_education)) %>% 
    scale() %>% 
    cor()
  

cor_pfas_adult <-
  pfas_adult %>%
    select(-bmi, -(seqn:hh_education)) %>% 
    scale() %>% 
    cor()


cor_pfas_adolescent %>%
  corrplot::corrplot(., type = "upper",  method = "ellipse",
                        tl.col = "black", tl.srt = 90, tl.cex = 0.7,
                        addCoef.col = "black", number.cex = 0.7)



cor_pfas_adult %>%
  corrplot::corrplot(., type = "upper",  method = "ellipse",
                        tl.col = "black", tl.srt = 90, tl.cex = 0.7,
                        addCoef.col = "black", number.cex = 0.7)


```

## Linear

```{r linear_map}

nest_lm_pfas_adolescent <-
  pfas_adolescent %>%
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(bmi ~ concentration + gender + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))


nest_lm_pfas_adult <-
  pfas_adult %>%
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(bmi ~ concentration + gender + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))

```


### PFAS beta coefficient estimate and model summary statistics
```{r lm_broom}
lm_pfas_coefs_adolescent <-
  nest_lm_pfas_adolescent %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    filter(term == "concentration") %>% 
    select(-term)

lm_pfas_coefs_adult <-
  nest_lm_pfas_adult %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    filter(term == "concentration") %>% 
    select(-term)




lm_all_coefs_adolescent <-
  nest_lm_pfas_adolescent %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    mutate(term = fct_inorder(term)) %>% 
    select(chemical, term, estimate) %>% 
    spread(key = term, value = estimate) 

lm_all_coefs_adult <-
  nest_lm_pfas_adult %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    mutate(term = fct_inorder(term)) %>% 
    select(chemical, term, estimate) %>% 
    spread(key = term, value = estimate) 



lm_model_summary_adolescent <-
nest_lm_pfas_adolescent %>% 
  mutate(models = map(models, broom::glance)) %>% 
  select(-data, -pred, -resids) %>% 
  unnest() 


lm_model_summary_adult <-
nest_lm_pfas_adult %>% 
  mutate(models = map(models, broom::glance)) %>% 
  select(-data, -pred, -resids) %>% 
  unnest() 


lm_pfas_coefs_adolescent %>%
  select(-statistic) %>% 
  left_join(lm_model_summary_adolescent, by = "chemical") %>%
  select(-r.squared, -sigma, -df, -p.value.y, -deviance, -df.residual) %>% 
  knitr::kable(digits = 3)


lm_pfas_coefs_adult %>%
  select(-statistic) %>% 
  left_join(lm_model_summary_adult, by = "chemical") %>%
  select(-r.squared, -sigma, -df, -p.value.y, -deviance, -df.residual) %>% 
  knitr::kable(digits = 3)

```

### Linear predicted values plots

```{r lm_plot}

nest_lm_pfas_adolescent %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth(method = lm) +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()


nest_lm_pfas_adult %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth(method = lm) +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()

```


## Penalized spline GAMs

k = 9 knots (limited by adolescent model)
```{r ps_pfas}
# have to reduce # of knows in adolescent models
#
nest_ps_pfas_adolescent <-
  pfas_adolescent %>%
  select(-pfdoa, -sb_pfoa) %>% 
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~gam(bmi ~ s(concentration, k = 9)  + gender + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))



nest_ps_pfas_adult <-
  pfas_adult %>%
  select(-pfdoa, -sb_pfoa) %>% 
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~gam(bmi ~ s(concentration, k = 9) + gender + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))

```


GAM penalized spline PFAS estimated degrees of freedom and model summary statistics
```{r ps_broom}
ps_edfs_adolescent <-
  nest_ps_pfas_adolescent %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    select(-term)


ps_edfs_adult <-
  nest_ps_pfas_adult %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    select(-term)

ps_summary_adolescent <-
  nest_ps_pfas_adolescent %>% 
    mutate(models = map(models, broom::glance)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() 

ps_summary_adult <-
  nest_ps_pfas_adult %>% 
    mutate(models = map(models, broom::glance)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() 

ps_edfs_adolescent %>% 
  left_join(ps_summary_adolescent, by = "chemical") %>% 
  select(-df, -deviance, -df.residual) %>%  
    knitr::kable(digits = 3)


ps_edfs_adult %>% 
  left_join(ps_summary_adolescent, by = "chemical") %>% 
  select(-df, -deviance, -df.residual) %>%  
    knitr::kable(digits = 3)

```

Penalized spline predicted values plots
```{r ps_plot}

nest_ps_pfas_adolescent %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()


nest_ps_pfas_adult %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()
```


## PCA Model

```{r pca_gam, warning = FALSE}
pca_scores <- read.csv("pca_scores.csv")

pfas_pca <-
  left_join(pfas, pca_scores, by = "seqn") 


pfas_pca_adolescent <-
  left_join(pfas, pca_scores, by = "seqn") %>% 
    filter(age <= 19)

pfas_pca_adult <-
  left_join(pfas, pca_scores, by = "seqn") %>% 
  filter(age > 19)
  
ps_pfas_pca_adolescent <-
  gam(bmi ~ s(PC1) + s(PC2) + s(PC3)  + gender + race_ethnicity + hh_education, data = pfas_pca_adolescent)

ps_pfas_pca_adult <-
  gam(bmi ~ s(PC1) + s(PC2) + s(PC3) + gender + race_ethnicity + hh_education, data = pfas_pca_adult)

```


```{r pca_broom}

ps_pfas_pca_adolescent %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)

ps_pfas_pca_adult %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)




ps_pfas_pca_adolescent %>% 
  broom::glance() %>% 
  knitr::kable(digits = 3)

ps_pfas_pca_adult %>% 
  broom::glance() %>% 
  knitr::kable(digits = 3)

```


Plot of PCA penalized splines with observed BMI points
```{r pca_plot, warning = FALSE}

pca_vars <- c("PC1", "PC2", "PC3")

map(pca_vars, function(x){
  p <- voxel::plotGAM(ps_pfas_pca_adolescent, smooth.cov = x) +
    geom_point(data = pfas_pca_adolescent, aes_string(y = "bmi", x = x ), alpha = 0.2, size = 0.5) +
    geom_rug(data = pfas_pca_adolescent, aes_string(y = "bmi", x = x ), alpha = 0.2)
  g <- ggplotGrob(p) 
  }) %>%
  {grid.arrange(grobs = (.), ncol = 3, nrow = 1)} 


map(pca_vars, function(x){
  p <- voxel::plotGAM(ps_pfas_pca_adult, smooth.cov = x) +
    geom_point(data = pfas_pca_adult, aes_string(y = "bmi", x = x ), alpha = 0.2, size = 0.5) +
    geom_rug(data = pfas_pca_adult, aes_string(y = "bmi", x = x ), alpha = 0.2)
  g <- ggplotGrob(p) 
  }) %>%
  {grid.arrange(grobs = (.), ncol = 3, nrow = 1)} 
```

