---
title: "rnhanes_nonlinear"
author: "SL"
date: "March 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
library(splines)

```

Load dataset from pipeline output
```{r load}
pfas <- read.csv("aamehs_data.csv") %>% 
   select(seqn, riagendr, ridageyr, ridreth3, indfmpir, pfdea, pfhxs, me_pfosa_acoh, pfna, pfua, pfdoa, n_pfoa, sb_pfoa, n_pfos, sm_pfos, bmxbmi) %>% 
  mutate(riagendr = factor(riagendr),
         ridreth3 = factor(ridreth3)) %>% 
  drop_na(bmxbmi)

glimpse(pfas)


#check for missing values
pfas %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) 
```

Missing values: 
indfmpir = 201  -- too high!, HH education level instead?
bmxbmi = 16

RIAGENDR - Gender
RIDAGEYR - Age in years at screening
RIDRETH3 - Race/Hispanic origin w/ NH Asian
DMQMILIZ - Served active duty in US Armed Forces

## Linear
Need an SES term --- finding negative association???
```{r pfoa_linear_model}
pfoa_linear <- lm(bmxbmi ~ n_pfoa + riagendr + ridageyr + ridreth3, 
              data = pfas)

summary(pfoa_linear)


pfoa_linear %>% 
  predict(., se.fit = TRUE, type = "terms" ) %>% 
  as.data.frame(.) %>% 
  mutate(pred = fit.n_pfoa) %>% 
  select(pred) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmxbmi)) %>% 
  ggplot(., aes(x = n_pfoa)) + 
    geom_line(aes(y = pred_bmi)) + 
    xlab("n_pfoa") + 
    ylab("Predicted BMI") 


```


```{r pfos_linear}
pfos_linear <- lm(bmxbmi ~ n_pfos + riagendr + ridageyr + ridreth3, 
              data = pfas)

summary(pfos_linear)
```


## Natural Spline Term 


```{r pfoa_natural_spline}

ns_pfoa <- lm(bmxbmi ~ ns(n_pfos, df = 3) + + riagendr + ridageyr + ridreth3, data = pfas)

summary(ns_pfoa)
AIC(ns_pfoa)

ns_pfoa %>% 
 predict(., se.fit = TRUE, type = "terms" ) %>% 
 as.data.frame(.) %>% 
  mutate(pred = fit.ns.n_pfos..df...3.,
         se = se.fit.ns.n_pfos..df...3.,
         lci = pred - 1.96*se,
         uci = pred + 1.96*se) %>%
  select(pred, se, lci, uci) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmxbmi),
         lci_bmi = lci + mean(bmxbmi),
         uci_bmi = uci + mean(bmxbmi)) %>% 
  ggplot(., aes(n_pfoa)) + 
      geom_line(aes(y = pred_bmi)) + 
      geom_line(aes(y = lci_bmi), color = "darkgrey") + 
      geom_line(aes(y = uci_bmi), color = "darkgrey") + 
      xlab("n_pfoa") + 
      ylab("Predicted BMI (95% CI)") +
      ylim(20,35)

```



















# 4L Create natural spline with gam()

mod.ns.gam.3 <- gam(aveBMI ~ ns(avePM.idw, df = 3) + maleunemp  + femaleunemp + ltHS + medhinc + medhval + 
                 per.black + per.latinx + per.asnam + climate_region, data = df)
termplot(mod.ns.gam.3, se = TRUE)


########################################
##### 5: Penalized Spline Term  ########
########################################
# 5a Construct model
# we can constuct penalized splines within gam() 
# default is 10 knots 
# only increase if need >~8.5 edf 
# (estimated degrees of freedom, vs. user-defined)
# penalty is estimated -- the model selects the penalty 
# that leads to the highest gcv

mod.ps <- gam(aveBMI ~ s(avePM.idw) + maleunemp  + femaleunemp + ltHS + medhinc + medhval + 
                per.black + per.latinx + per.asnam + climate_region, data = df)

# 5b Model Summary 

summary(mod.ps)

# 5c Extract Penalty 
# this is the penalty estimated by the model 

mod.ps$sp

# 5d Model Plot 
# plot.gam offers a nice default plot
# a quick way to plot the change in predicted aveBMI with avePM.idw

plot(mod.ps)

# 5e Plot with ggplot 
# we can also recreate this plot with ggplot

# 5e construct predictions based on the model

predBMI.ps <- predict(mod.ps, se.fit = TRUE, type = "terms" )

# 5f convert to dataframe 

predBMI.ps <- as.data.frame(predBMI.ps)

# 5g Combine predictions and standard errors
# it just has different names since its one term 

predBMI.ps <- predBMI.ps %>% 
  mutate( pred = fit.s.avePM.idw.,
          se = se.fit.s.avePM.idw.)

# 5h Compute 95% confidence intervals 

predBMI.ps <- predBMI.ps %>% 
  mutate( lci = pred - 1.96*se,
          uci = pred + 1.96*se)

# 5i Keep only variables we need 

predBMI.ps2 <- predBMI.ps %>% select(pred, se, lci, uci)

# 5j Combine with data 

predBMI.ps2 <- predBMI.ps2 %>% bind_cols(df)

# 5k Uncenter data 

predBMI.ps2 <- predBMI.ps2 %>% mutate(predBMI = pred + mean(aveBMI),
                                      lciBMI = lci + mean(aveBMI),
                                      uciBMI = uci + mean(aveBMI))
# 5l Plot

ggplot(predBMI.ps2, aes(avePM.idw)) + 
  geom_line(aes(y = predBMI)) + 
  geom_line(aes(y = lciBMI), color = "darkgrey") + 
  geom_line(aes(y = uciBMI), color = "darkgrey") + 
  xlab(expression("Average Annual PM"[2.5])) + 
  ylab("Predicted BMI (95% CI)") + 
  ylim(26, 29)

#########################################################
##### Footnote: Plotting All the Models Together ########
#########################################################
# create a column of TermName for each dataframe of predictions 
# we will use this column to keep track of which model 
# the predictions came from 

# A Create linear model 
predBMI.lin  <- predict(mod.lin, se.fit = TRUE, type = "terms" )
predBMI.lin  <- as.data.frame(predBMI.lin)
predBMI.lin  <- predBMI.lin %>% mutate(pred = fit.avePM.idw)
predBMI.lin2 <- predBMI.lin %>% select(pred)
predBMI.lin2 <- predBMI.lin2 %>% bind_cols(df)
predBMI.lin2 <- predBMI.lin2 %>% mutate(predBMI = pred + mean(aveBMI))

# B Create a column with the model name 
# we will use this to keep track of the models once we combine them
predBMI.lin3  <- predBMI.lin2  %>% mutate(ModelName = "Linear")
predBMI.quad3 <- predBMI.quad2 %>% mutate(ModelName = "Quadratic Term")
predBMI.pls3  <- predBMI.pls2  %>% mutate(ModelName = "Piecewise Linear Spline")
predBMI.ns.3.3<- predBMI.ns.3.2%>% mutate(ModelName = "Natural Spline 3 df")
predBMI.ps3   <- predBMI.ps2   %>% mutate(ModelName = "Penalized Spline")

# C Combine the predictions 

predBMI.tot <- bind_rows(predBMI.lin3,
                         predBMI.quad3, 
                         predBMI.pls3,
                         predBMI.ns.3.3,
                         predBMI.ps3)
# D Plot!

all.models.plot <-  ggplot(predBMI.tot, aes(avePM.idw)) + 
  geom_line(aes(y = predBMI, color = ModelName)) + 
  xlab(expression("Average Annual PM"[2.5])) + 
  ylab("Predicted BMI") + 
  ylim(26, 29)

all.models.plot 

# save plot as a pdf
pdf(paste0(OutputPath, "all_models_plot.pdf"), width = 10)
all.models.plot
dev.off()

###################################
##### Footnote: Cubic Term ########
###################################
# A Create model 
# we can create new terms within the model statement 
# using the I() command 

mod.cub <- lm(aveBMI ~ avePM.idw + I(avePM.idw^2) + I(avePM.idw^3) + 
              maleunemp  + femaleunemp + ltHS + medhinc + medhval +
              per.black + per.latinx + per.asnam + climate_region, 
              data = df)

# B Model Summary 

summary(mod.cub)

# C Construct predictions based on the model

predBMI <- predict(mod.cub, se.fit = TRUE, type = "terms" )

# D convert to dataframe 

predBMI <- as.data.frame(predBMI)

# E Combine predictions and standard errors

predBMI <- predBMI %>% 
  mutate( pred = fit.avePM.idw + fit.I.avePM.idw.2.+ fit.I.avePM.idw.3.,
          se = se.fit.avePM.idw + se.fit.I.avePM.idw.2.+ se.fit.I.avePM.idw.3.)

# F Compute 95% confidence intervals 

predBMI <- predBMI %>% 
  mutate( lci = pred - 1.96*se,
          uci = pred + 1.96*se)

# G Keep only variables we need 

predBMI2 <- predBMI %>% select(pred, se, lci, uci)

# H Combine with data 

predBMI2 <- predBMI2 %>% bind_cols(df)

# I Plot

ggplot(predBMI2, aes(avePM.idw)) + 
  geom_line(aes(y = pred)) + 
  geom_line(aes(y = lci), color = "darkgrey") + 
  geom_line(aes(y = uci), color = "darkgrey") + 
  xlab(expression("Average Annual PM"[2.5])) + 
  ylab("Predicted aveBMI")

#######################################
##### Footnote: Multiple Plots ########
#######################################
# A Create predictions and 95% CI for 2 df
# A1 Create predictions for 2 degrees of freedom
predBMI <- predict(mod.ns.2, se.fit = TRUE, type = "terms" )
# A2Convert to dataframe 
predBMI <- as.data.frame(predBMI)
# A3 Combine predictions and standard errors
predBMI <- predBMI %>% 
  mutate( pred = fit.ns.avePM.idw..df...2.,
          se = se.fit.ns.avePM.idw..df...2.)
# A4 Compute 95% confidence intervals 
predBMI <- predBMI %>% 
  mutate( lci = pred - 1.96*se,
          uci = pred + 1.96*se)
# A5 Keep only variables we need 
predBMI2 <- predBMI %>% select(pred, se, lci, uci) %>% 
  mutate(Model = "2 df")
# A6 Combine with data 
predBMI.2 <- predBMI2 %>% bind_cols(df)

# B Create predictions and 95% CI for 3 df
# B1 Create predictions for 3 degrees of freedom
predBMI <- predict(mod.ns.3, se.fit = TRUE, type = "terms" )
# B2 convert to dataframe 
predBMI <- as.data.frame(predBMI)
# B3 Combine predictions and standard errors
predBMI <- predBMI %>% 
  mutate( pred = fit.ns.avePM.idw..df...3.,
          se = se.fit.ns.avePM.idw..df...3.)
# B4 Compute 95% confidence intervals 
predBMI <- predBMI %>% 
  mutate( lci = pred - 1.96*se,
          uci = pred + 1.96*se)
# B5 Keep only variables we need
predBMI2 <- predBMI %>% select(pred, se, lci, uci)%>% 
  mutate(Model = "3 df")
# B6 Combine with data 
predBMI.3 <- predBMI2 %>% bind_cols(df)

# C Create predictions and 95% CI for 4 df
# C1 Create predictions for 4 degrees of freedom
predBMI <- predict(mod.ns.4, se.fit = TRUE, type = "terms" )
# C2 convert to dataframe 
predBMI <- as.data.frame(predBMI)
# C3 Combine predictions and standard errors
predBMI <- predBMI %>% 
  mutate( pred = fit.ns.avePM.idw..df...4.,
          se = se.fit.ns.avePM.idw..df...4.)
# C4 Compute 95% confidence intervals 
predBMI <- predBMI %>% 
  mutate( lci = pred - 1.96*se,
          uci = pred + 1.96*se)
# C5 Keep only variables we need
predBMI2 <- predBMI %>% select(pred, se, lci, uci) %>% 
  mutate(Model = "4 df")
# C6 Combine with data 
predBMI.4 <- predBMI2 %>% bind_cols(df)
# C7 combine all the data 
allModels <- bind_rows(predBMI.2, predBMI.3, predBMI.4)

# D Uncenter the data - add mean aveBMI
allModels <- allModels %>% mutate(predBMI = pred + mean(aveBMI),
                                  lciBMI = lci + mean(aveBMI),
                                  uciBMI = uci + mean(aveBMI))

# E Plot 

pdf(paste0(OutputPath, "three_ns_model_plot.pdf"))

# no confidence intervals
ggplot(allModels, aes(avePM.idw)) + 
  geom_line(aes(y = predBMI, color = Model)) + 
  xlab(expression("Average Annual PM"[2.5])) + 
  ylab("Predicted BMI") + 
  ylim(26, 29)

# confidence intervals
ggplot(allModels, aes(avePM.idw)) + 
  geom_line(aes(y = predBMI, color = Model)) + 
  geom_line(aes(y = lciBMI, color = Model), alpha = 0.1) + 
  geom_line(aes(y = uciBMI, color = Model), alpha = 0.1) + 
  xlab(expression("Average Annual PM"[2.5])) + 
  ylab("Predicted BMI (95% CI)") + 
  ylim(26, 29)

dev.off()







