---
title: "rnhanes_nonlinear_final"
author: "SL"
date: "April 17, 2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
library(splines)
library(ggcorrplot)
library(corrplot)
library(modelr)

```

Load dataset from pipeline output
```{r load}
pfas <- read.csv("aamehs_data.csv") %>% 
   select(seqn, gender, age, race_ethnicity, hh_education, pfdea, pfhxs, me_pfosa_acoh, pfna, pfua, pfdoa, n_pfoa, sb_pfoa, n_pfos, sm_pfos, bmi) %>% 
  mutate(gender = factor(gender),
         race_ethnicity = factor(race_ethnicity),
         hh_education = factor(hh_education)) %>% 
  drop_na()


```


## Tables
```{r, warning = FALSE}

pfas %>% 
  select(gender, age, race_ethnicity, hh_education, bmi, everything(), -seqn) %>% 
  summary(.) 

```


## Plots
```{r}
pfas %>%
    select(bmi, pfdea:sm_pfos) %>%
    gather(pfdea:sm_pfos, key = "variable", value = "value") %>% 
  ggplot(aes(value, bmi)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~variable) +
    theme_bw()

```

## Correlations

```{r}
cor_pfas <-
  pfas %>% 
    select(-bmi, -(seqn:hh_education)) %>% 
    scale() %>% 
    cor()
  


cor_pfas %>%
  corrplot::corrplot(., type = "upper",  method = "ellipse",
                        tl.col = "black", tl.srt = 90, tl.cex = 0.7,
                        addCoef.col = "black", number.cex = 0.7)



```



## Linear



### Linear model
```{r linear_map}

nest_lm_pfas <-
  pfas %>%
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(bmi ~ concentration + gender + age + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))


nest_lm_pfas %>% 
  mutate(models = map(models, broom::tidy)) %>% 
  select(-data, -pred, -resids) %>% 
  unnest() %>%  
  select(chemical, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  spread(key = term, value = estimate) %>% 
  knitr::kable(digits = 3)

nest_lm_pfas %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth(method = lm) +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()



```


## Natural Spline Term 

### glm natural spline
```{r pfoa_natural_spline}

ns_pfoa <- lm(bmi ~ ns(n_pfoa, df = 3) + gender + age + race_ethnicity + hh_education, data = pfas)

summary(ns_pfoa)
AIC(ns_pfoa)

ns_pfoa %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)


plot(ns_pfoa)

ns_pfoa %>% 
 predict(., se.fit = TRUE, type = "terms" ) %>% 
 as.data.frame(.) %>% 
  mutate(pred = fit.ns.n_pfoa..df...3.,
         se = se.fit.ns.n_pfoa..df...3.,
         lci = pred - 1.96*se,
         uci = pred + 1.96*se) %>%
  select(pred, se, lci, uci) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmi),
         lci_bmi = lci + mean(bmi),
         uci_bmi = uci + mean(bmi)) %>% 
  ggplot(., aes(n_pfoa)) + 
      geom_line(aes(y = pred_bmi)) + 
      geom_line(aes(y = lci_bmi), color = "darkgrey") + 
      geom_line(aes(y = uci_bmi), color = "darkgrey") + 
      xlab("n_pfoa") + 
      ylab("Predicted BMI (95% CI)") +
      ylim(20,35)


```


### gam natural spline



```{r gam_ns}
pfoa_gam_ns <- gam(bmi ~ ns(n_pfoa, df = 3) + gender + age + race_ethnicity + hh_education, data = pfas)


summary(pfoa_gam_ns)

pfoa_gam_ns %>% 
 predict(., se.fit = TRUE, type = "terms" ) %>% 
 as.data.frame(.) %>% 
  mutate(pred = fit.ns.n_pfoa..df...3.,
         se = se.fit.ns.n_pfoa..df...3.,
         lci = pred - 1.96*se,
         uci = pred + 1.96*se) %>%
  select(pred, se, lci, uci) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmi),
         lci_bmi = lci + mean(bmi),
         uci_bmi = uci + mean(bmi)) %>% 
  ggplot(., aes(n_pfoa)) + 
      geom_line(aes(y = pred_bmi)) + 
      geom_line(aes(y = lci_bmi), color = "darkgrey") + 
      geom_line(aes(y = uci_bmi), color = "darkgrey") + 
      xlab("n_pfoa") + 
      ylab("Predicted BMI (95% CI)") +
      ylim(20,35)




```
```


## Penalized spline

```{r ps_pfoa}
ps_pfoa <- gam(bmi ~ s(n_pfoa) + gender + age + race_ethnicity + hh_education, data = pfas)

summary(ps_pfoa)
ps_pfoa$sp   # extract Penalty 
plot(ps_pfoa)

ps_pfoa %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)



ps_pfoa %>% 
 predict(., se.fit = TRUE, type = "terms" ) %>% 
 as.data.frame(.) %>% 
  mutate(pred = fit.s.n_pfoa.,
         se = se.fit.s.n_pfoa.,
         lci = pred - 1.96*se,
         uci = pred + 1.96*se) %>%
  select(pred, se, lci, uci) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmi),
         lci_bmi = lci + mean(bmi),
         uci_bmi = uci + mean(bmi)) %>% 
  ggplot(., aes(n_pfoa)) + 
      geom_line(aes(y = pred_bmi)) + 
      geom_line(aes(y = lci_bmi), color = "darkgrey") + 
      geom_line(aes(y = uci_bmi), color = "darkgrey") + 
      xlab("n_pfoa") + 
      ylab("Predicted BMI (95% CI)") +
      ylim(20,35)







```


## PCA Loading Models


```{r}
pca_scores <- read.csv("pca_scores.csv")

pfas_pca <-
  left_join(pfas, pca_scores, by = "seqn") 
  
ps_pfas_pca <-
  gam(bmi ~ s(PC1) + s(PC2) + s(PC3) + gender + age + race_ethnicity + hh_education, data = pfas_pca)


summary(ps_pfas_pca)
ps_pfas_pca$sp   # extract Penalty 
plot(ps_pfas_pca)

ps_pfas_pca %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)



ps_pfoa %>% 
 predict(., se.fit = TRUE, type = "terms" ) %>% 
 as.data.frame(.) %>% 
  mutate(pred = fit.s.n_pfoa.,
         se = se.fit.s.n_pfoa.,
         lci = pred - 1.96*se,
         uci = pred + 1.96*se) %>%
  select(pred, se, lci, uci) %>% 
  bind_cols(pfas) %>% 
  mutate(pred_bmi = pred + mean(bmi),
         lci_bmi = lci + mean(bmi),
         uci_bmi = uci + mean(bmi)) %>% 
  ggplot(., aes(n_pfoa)) + 
      geom_line(aes(y = pred_bmi)) + 
      geom_line(aes(y = lci_bmi), color = "darkgrey") + 
      geom_line(aes(y = uci_bmi), color = "darkgrey") + 
      xlab("n_pfoa") + 
      ylab("Predicted BMI (95% CI)") +
      ylim(20,35)
```


## Compare Models

```{r compare_models, eval = FALSE}

broom::glance(pfoa_linear) %>% 
  bind_rows(broom::glance(ns_pfoa)) %>% 
  bind_rows(broom::glance(ps_pfoa)) %>% 
  bind_rows(broom::glance(ps_pfas_pca)) %>% 
  mutate(model = c("pfoa_linear", "pfoa_ns", "pfoa_ps", "pca3_ps")) %>% 
  select(model, everything()) %>% 
  knitr::kable()
              
            
```

