---
title: "rnhanes_nonlinear_final"
author: "SL"
date: "April 17, 2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, mesage = FALSE, warning = FALSE)
library(tidyverse)
library(mgcv)
library(splines)
library(ggcorrplot)
library(corrplot)
library(modelr)
library(voxel)
library(gridExtra)

```

Load dataset from pipeline output
```{r load}
pfas <- read.csv("aamehs_data.csv") %>% 
   select(seqn, gender, age, race_ethnicity, hh_education, pfdea, pfhxs, me_pfosa_acoh, pfna, pfua, pfdoa, n_pfoa, sb_pfoa, n_pfos, sm_pfos, bmi) %>% 
  mutate(gender = factor(gender),
         race_ethnicity = factor(race_ethnicity),
         hh_education = factor(hh_education)) %>% 
  drop_na()


```


## Data Summary
```{r, warning = FALSE}

pfas %>% 
  select(gender, age, race_ethnicity, hh_education, bmi, everything(), -seqn) %>% 
  summary(.) 

```


## Plots
```{r}
pfas %>%
    select(bmi, pfdea:sm_pfos) %>%
    gather(pfdea:sm_pfos, key = "variable", value = "value") %>% 
  ggplot(aes(value, bmi)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~variable) +
    theme_bw()

```

## Correlations

```{r}
cor_pfas <-
  pfas %>% 
    select(-bmi, -(seqn:hh_education)) %>% 
    scale() %>% 
    cor()
  


cor_pfas %>%
  corrplot::corrplot(., type = "upper",  method = "ellipse",
                        tl.col = "black", tl.srt = 90, tl.cex = 0.7,
                        addCoef.col = "black", number.cex = 0.7)



```

## Linear

```{r linear_map}

nest_lm_pfas <-
  pfas %>%
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(bmi ~ concentration + gender + age + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))


```


### PFAS beta coefficient estimate and model summary statistics
```{r lm_broom}
lm_pfas_coefs <-
  nest_lm_pfas %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    filter(term == "concentration") %>% 
    select(-term)


lm_all_coefs <-
  nest_lm_pfas %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    mutate(term = fct_inorder(term)) %>% 
    select(chemical, term, estimate) %>% 
    spread(key = term, value = estimate) 

lm_model_summary <-
nest_lm_pfas %>% 
  mutate(models = map(models, broom::glance)) %>% 
  select(-data, -pred, -resids) %>% 
  unnest() 


lm_pfas_coefs %>%
  select(-statistic) %>% 
  left_join(lm_model_summary, by = "chemical") %>%
  select(-r.squared, -sigma, -df, -p.value.y, -deviance, -df.residual) %>% 
  knitr::kable(digits = 3)

```

### Linear predicted values plots

```{r lm_plot}

nest_lm_pfas %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth(method = lm) +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()
```


## Penalized spline GAMs

```{r ps_pfas}

nest_ps_pfas <-
  pfas %>%
  select(-pfdoa, -sb_pfoa) %>% 
  gather(pfdea:sm_pfos, key = "chemical", value = "concentration") %>% 
  group_by(chemical) %>% 
  nest() %>% 
  mutate(models = map(data, ~gam(bmi ~ s(concentration) + gender + age + race_ethnicity + hh_education, data = .x)),
         pred  = map2(data, models, modelr::add_predictions),
         resids = map2(data, models, add_residuals))


```


GAM penalized spline PFAS estimated degrees of freedom and model summary statistics
```{r ps_broom}
ps_edfs <-
  nest_ps_pfas %>% 
    mutate(models = map(models, broom::tidy)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() %>%  
    select(-term)

ps_summary <-
  nest_ps_pfas %>% 
    mutate(models = map(models, broom::glance)) %>% 
    select(-data, -pred, -resids) %>% 
    unnest() 

ps_edfs %>% 
  left_join(ps_summary, by = "chemical") %>% 
  select(-df, -deviance, -df.residual) %>%  
    knitr::kable(digits = 3)

```

# Penalized spline predicted values plots

```{r ps_plot}

nest_ps_pfas %>% 
  unnest(pred) %>% 
    ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~chemical) +
    scale_x_log10() +
    theme_bw()


gam_plot <-
nest_ps_pfas %>%
  mutate(chemical = recode(chemical, "n_pfoa" = "PFOA"),
         chemical = recode(chemical, "n_pfos" = "PFOS")) %>% 
    unnest(pred) %>% 
   filter(chemical == c("PFOA", "PFOS")) %>%
       ggplot(aes(concentration, pred)) +
      geom_point(size = 0.5) +
    geom_smooth() +
    facet_wrap(~chemical) +
    xlab("Serum Concentration (ng/mL)") +
    ylab("Predicted BMI") +
    theme_bw()


ggsave("gam_plot.png", plot = gam_plot, scale = 1, dpi = 320)
  
  





```


## PCA Model

```{r pca_gam, warning = FALSE}
pca_scores <- read.csv("pca_scores.csv")

pfas_pca <-
  left_join(pfas, pca_scores, by = "seqn") 
  
ps_pfas_pca <-
  gam(bmi ~ s(PC1) + s(PC2) + s(PC3) + gender + age + race_ethnicity + hh_education, data = pfas_pca)

summary(ps_pfas_pca)

```


```{r pca_broom}

ps_pfas_pca %>% 
  broom::tidy() %>% 
  select(-statistic) %>% 
  knitr::kable(digits = 3)

ps_pfas_pca %>% 
  broom::glance() %>% 
  knitr::kable(digits = 3)

```


Plot of PCA penalized splines with observed BMI points
```{r pca_plot, warning = FALSE}

pca_vars <- c("PC1", "PC2", "PC3")

map(pca_vars, function(x){
  p <- voxel::plotGAM(ps_pfas_pca, smooth.cov = x) +
    geom_point(data = pfas_pca, aes_string(y = "bmi", x = x ), alpha = 0.2, size = 0.5) +
    geom_rug(data = pfas_pca, aes_string(y = "bmi", x = x ), alpha = 0.2)
  g <- ggplotGrob(p) 
  }) %>%
  {grid.arrange(grobs = (.), ncol = 3, nrow = 1)} 


pca_gam_plot <-
map(pca_vars, function(x){
  p <- voxel::plotGAM(ps_pfas_pca, smooth.cov = x) +
    geom_point(data = pfas_pca, aes_string(y = "bmi", x = x ), alpha = 0.2, size = 0.5) +
    geom_rug(data = pfas_pca, aes_string(y = "bmi", x = x ), alpha = 0.2)+
    xlab("Principal Component") +
    ylab("Predicted BMI") +
    ggtitle("")

  g <- ggplotGrob(p) 
  }) %>%
  {grid.arrange(grobs = (.), ncol = 3, nrow = 1)} 


```



PLots for paper

PCA patchwork


```{r PCA, message=FALSE}

aamehs_data <- read_csv("aamehs_data.csv")

pca_data <- aamehs_data %>% 
  select(pfdea, pfhxs, me_pfosa_acoh, pfna, pfua, pfdoa, n_pfoa, sb_pfoa, n_pfos, sm_pfos)
library(psych)
pca <- prcomp(pca_data, scale. = TRUE) 

ls(pca)
summary(pca)

eigenvalues.v <- pca$sdev^2
perc_variance.v <- eigenvalues.v/sum(eigenvalues.v)
perc_variance.v <- round(100 * perc_variance.v, 1)
cumulative_perc_var.v <- cumsum(perc_variance.v)
eigenvalues.df <- data.frame(1:length(eigenvalues.v), eigenvalues.v, perc_variance.v, cumulative_perc_var.v)
colnames(eigenvalues.df) <- c("Principal Component", "Eigenvalues", 
                              "Proportion Var Explained", "Cumulative Proportion")
# Scree Plot
library(factoextra)
fviz_eig(pca, main = "Percent Variance Explained \n by Principal Component",
         xlab = "Principal component",
         ylim = c(0,70)) 

# Visualize Loadings
loadings.df <- as.data.frame.matrix(pca$rotation) 
loadings.df$Constituent <- row.names(loadings.df)
loadings.long <- loadings.df %>% 
  gather(key = "PC", value = "Loading", -Constituent) 

ggplot(loadings.long, aes(x = Constituent, y = Loading)) + 
  geom_col() +                             # creates a column for each loading
  geom_hline(yintercept = 0, size = 0.2) + # creates a line at 0
  facet_wrap(~ PC) +                       # creates a distinct box for each PC 
  theme_bw() +                             # sets theme options
  theme(strip.background = element_rect(fill = "white")) +
  labs(x = expression("PFOS Constituents"),
       y = "Loadings") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

loadings_plot <-
  loadings.long %>% 
 mutate(PC = factor(PC, levels = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10"))) %>% 
  filter(PC %in% c("PC1", "PC2", "PC3")) %>% 
ggplot(aes(x = Constituent, y = Loading)) + 
  geom_col() +                             # creates a column for each loading
  geom_hline(yintercept = 0, size = 0.2) + # creates a line at 0
  facet_wrap(~ PC) +                       # creates a distinct box for each PC 
  theme_bw() +                             # sets theme options
  theme(strip.background = element_rect(fill = "white")) +
  labs(x = expression("PFOS Constituents"),
       y = "Loadings") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


```{r pca_patchwork}

library(patchwork)
loadings_plot
pca_gam_plot

pca_fig <-
loadings_plot / pca_gam_plot

ggsave("pca_fig.png", plot = pca_fig, scale = 1.5, dpi = 320)



```




