---
title: "summary_checkin"
author: "Team LSD: Adnan, Jenni, Stephen"
date: "February 13, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Per- and polyfluoroalkyl substances (PFAS) and Body Mass
## 26 February 2019 Check-In



### 1. Water and food consumption and consumer product use influence PFAS serum concentrations.

**IV**: Survey Questions: drinking water sources, drinking water consumption, canned goods, microwave popcorn, beauty and personal care products (including dental floss)

**DV**: Serum PFAS concentrations (Outcome): PFDeA, PFHxS, Me-PFOSA-AcOH, PFNA, PFUA, PFDoA, n-PFOA, Sb-PFOA, n-PFOS, Sm-PFOS

### 2. Elevated PFAS concentrations contribute to higher BMI.

**IV**: Serum PFAS concentrations (Outcome): PFDeA, PFHxS, Me-PFOSA-AcOH, PFNA, PFUA, PFDoA, n-PFOA, Sb-PFOA, n-PFOS, Sm-PFOS

**DV**: Body weight, BMI

Potential confounders, effect modifiers, or co-variates of interest: 
* age 
* gestational diabetes
* pre-diabetes / diabetes
* sex 
* smoking status
* household income
* alcohol
* waist circumference, 2 years of age and older
* sagittal abdominal diameter, 8 years of age and older

*References*: https://www.ncbi.nlm.nih.gov/pubmed/?term=PFAS+and+BMI 

### 3. Serum PFAS levels serve as a mediator between water/food consumption and body mass.

**DAG**:
  
water/food/product use (exposure) --> PFAS serum concentrations (internal dose) --> increased body mass (biological effect)  


```{r packages, message = FALSE}

library(tidyverse)
library(RNHANES)
library(survey)

```


## NHANES Dietary Interviews

*Reference*: https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1IFF_I.htm

Detailed information about each food/beverage item (including the description, amount of, and nutrient content) reported by each participant is included in the Individual Foods files.

* Drinking water variables:

    + Total plain water drank yesterday (gm)
    + Total tap water drank yesterday (gm)
    + Total bottled water drank yesterday (gm)
    + Tap water source

* Fish consumption during past 30 days (by type) 

## PFAS

### Load and Inspect PFAS data
NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/PFAS_H.htm
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PFAS_I.htm

```{r load_pfas, message = FALSE}

 
pfas_load <- nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE)

pfas_survey <- nhanes_survey_design(pfas_load, "WTSB2YR")

pfas_data <- nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE) %>% 
  janitor::clean_names() %>% 
  rename(pfdea = lbxpfde, pfhxs = lbxpfhs, me_pfosa_acoh = lbxmpah,	pfna = lbxpfna, pfua = lbxpfua,	pfdoa = lbxpfdo, n_pfoa = lbxnfoa,	sb_pfoa = lbxbfoa, n_pfos = lbxnfos,	sm_pfos = lbxmfos)


```

### 2015-2016 PFAS summary table


```{r summary_tables_remove_na}

#pfas_summary <-
#  pfas_data %>%
#    select(seqn, pfdea, pfhxs, me_pfosa_acoh,	pfna, pfua,	pfdoa, n_pfoa, sb_pfoa, n_pfos,	sm_pfos) %>% 
#    gather(key = "analyte", value = "value", pfdea:sm_pfos) %>% 
#    group_by(analyte) %>% 
#    na.omit() %>% 
#    summarise(n = n(), 
#              mean = mean(value),
#              sd = sd(value))


pfas_summary <-
  pfas_load %>%
    select(SEQN, LBXPFDE, LBXPFHS, LBXMPAH, LBXPFNA, LBXPFUA, LBXPFDO, LBXNFOA, LBXBFOA, LBXNFOS, LBXMFOS) %>% 
    gather(key = "variable", value = "value", LBXPFDE:LBXMFOS) %>% 
    group_by(variable) %>% 
    na.omit() %>% 
    summarise(n = n(), 
              mean = mean(value),
              sd = sd(value))


# Weighted

pfas_inputs <- as.data.frame(matrix(c(
  # CYCLE        COLUMN    COMMENT     WEIGHTS
   "2015-2016", "LBXPFDE", "LBDPFDEL", "WTSB2YR",
   "2015-2016", "LBXPFHS", "LBDPFHSL", "WTSB2YR",
   "2015-2016", "LBXMPAH", "LBDMPAHL", "WTSB2YR",
   "2015-2016", "LBXPFNA", "LBDPFNAL", "WTSB2YR",
   "2015-2016", "LBXPFUA", "LBDPFUAL", "WTSB2YR",
   "2015-2016", "LBXPFDO", "LBDPFDOL", "WTSB2YR",
   "2015-2016", "LBXNFOA", "LBDNFOAL", "WTSB2YR",
   "2015-2016", "LBXBFOA", "LBDBFOAL", "WTSB2YR",
   "2015-2016", "LBXNFOS", "LBDNFOSL", "WTSB2YR",
   "2015-2016", "LBXMFOS", "LBDMFOSL", "WTSB2YR"
), ncol = 4, byrow = TRUE), stringsAsFactors = FALSE)

names(pfas_inputs) <- c("cycle", "column", "comment_column", "weights_column")

weighted_median <-
  nhanes_quantile(pfas_load, pfas_inputs, quantiles = 0.5) %>% 
  select(column, weighted_median = value)

weighted_mean <-
  nhanes_survey(svymean, pfas_load, pfas_inputs, na.rm = TRUE) %>% 
    select(column, weighted_mean = value)




codes <- data.frame(
  variable = c("LBXPFDE", "LBXPFHS", "LBXMPAH", "LBXPFNA", "LBXPFUA", "LBXPFDO", "LBXNFOA", "LBXBFOA", "LBXNFOS", "LBXMFOS"),
  analyte = c("pfdea", "pfhxs", "me_pfosa_acoh",	"pfna", "pfua",	"pfdoa", "n_pfoa", "sb_pfoa", "n_pfos",	"sm_pfos"))


pfas_table <-
  pfas_summary %>% 
    left_join(weighted_mean, by = c("variable" = "column")) %>% 
    left_join(weighted_median, by = c("variable" = "column")) %>% 
    left_join(codes, by = "variable") %>% 
    subset(select = c(1, 7, 2, 3, 4, 5, 6))

pfas_table %>%
  knitr::kable(digits = 2) 

kable_as_image(kable_input, filename = NULL, file_format = "png",
  latex_header_includes = NULL, keep_pdf = FALSE, density = 300,
  keep_tex = FALSE)

```

## Body Mass

NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BMX_I.htm

### Load and Inspect BMI data

```{r load body mass data, message = FALSE}
bodymass_data <- nhanes_load_data("BMX_I", "2015-2016", demographics = TRUE)
as_tibble(bodymass_data)
```

### Inspect body mass data from 2015-2016
```{r examine_bmi, warning = FALSE, message = FALSE}
# bodymass_data %>%  nhanes_detection_frequency("BMXBMI", "BMXBMI", "WTMEC2YR") # not completely sure on weight
bodymass_data %>% nhanes_sample_size("BMXBMI", "BMXBMI", "WTMEC2YR")
bodymass_data %>%  nhanes_quantile("BMXBMI","BMXBMI", "WTMEC2YR", quantiles = c(0.5, 0.95))
```

## Merged Dataset

```{r}

pfas_data_clean = nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE) %>% 
  janitor::clean_names() %>% 
  select(seqn, cycle, sddsrvyr, riagendr, ridageyr, ridreth3, dmdeduc3, dmdeduc2, wtint2yr, wtmec2yr, lbxpfde:lbdmfosl) %>% 
  rename(pfdea = lbxpfde, pfhxs = lbxpfhs, me_pfosa_acoh = lbxmpah,	pfna = lbxpfna, pfua = lbxpfua,	pfdoa = lbxpfdo, n_pfoa = lbxnfoa,	sb_pfoa = lbxbfoa, n_pfos = lbxnfos,	sm_pfos = lbxmfos)

bodymass_data_clean = bodymass_data %>% 
  janitor::clean_names() %>% 
  select(seqn, bmxbmi, bmxwt, bmiwt)

pfas_bodymass_clean = left_join(pfas_data_clean, bodymass_data_clean, by = "seqn")

```

