---
title: "summary_checkin"
author: "Team LSD: Adnan, Jenni, Stephen"
date: "February 13, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Per- and polyfluoroalkyl substances (PFAS) and Body Mass
## 26 February 2019 Check-In



### 1. Water and food consumption and consumer product use influence PFAS serum concentrations.

**IV**: Survey Questions: drinking water sources, drinking water consumption, canned goods, microwave popcorn, beauty and personal care products (including dental floss)

**DV**: Serum PFAS concentrations (Outcome): PFDeA, PFHxS, Me-PFOSA-AcOH, PFNA, PFUA, PFDoA, n-PFOA, Sb-PFOA, n-PFOS, Sm-PFOS

### 2. Elevated PFAS concentrations contribute to higher BMI.

**IV**: Serum PFAS concentrations (Outcome): PFDeA, PFHxS, Me-PFOSA-AcOH, PFNA, PFUA, PFDoA, n-PFOA, Sb-PFOA, n-PFOS, Sm-PFOS

**DV**: Body weight, BMI

Potential confounders, effect modifiers, or co-variates of interest: 
* age 
* gestational diabetes
* pre-diabetes / diabetes
* sex 
* smoking status
* household income
* alcohol
* waist circumference, 2 years of age and older
* sagittal abdominal diameter, 8 years of age and older

*References*: https://www.ncbi.nlm.nih.gov/pubmed/?term=PFAS+and+BMI 

### 3. Serum PFAS levels serve as a mediator between water/food consumption and body mass.

**DAG**:
  
water/food/product use (exposure) --> PFAS serum concentrations (internal dose) --> increased body mass (biological effect)  


```{r packages, message = FALSE}
library(tidyverse)
library(RNHANES)
library(survey)
```


NHANES files and the comprehensive variable list
```{r load_file_names, cache = TRUE, message = FALSE}
files <- nhanes_data_files()  #1,328 obs of 10 variables
variables <- nhanes_variables()  #48,398 obs of 10 variables

```

```{r explore_files}
files %>% count(cycle) %>% View
files %>% count(component)
```


## NHANES Dietary Interviews

*Reference*: https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DR1IFF_I.htm

Detailed information about each food/beverage item (including the description, amount of, and nutrient content) reported by each participant is included in the Individual Foods files.

* Drinking water variables:

    + Total plain water drank yesterday (gm)
    + Total tap water drank yesterday (gm)
    + Total bottled water drank yesterday (gm)
    + Tap water source

* Fish consumption during past 30 days (by type) 

## PFAS

### Load and Inspect PFAS data
NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/PFAS_H.htm
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PFAS_I.htm

```{r load_pfas, message = FALSE}

 
pfas_data <- nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE)

pfas_survey <- nhanes_survey_design(pfas_data, "WTSB2YR")

pfas_data <- nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE) %>% 
  janitor::clean_names() %>% 
  rename(pfdea = lbxpfde, pfhxs = lbxpfhs, me_pfosa_acoh = lbxmpah,	pfna = lbxpfna, pfua = lbxpfua,	pfdoa = lbxpfdo, n_pfoa = lbxnfoa,	sb_pfoa = lbxbfoa, n_pfos = lbxnfos,	sm_pfos = lbxmfos)


```

### 2015-2016 PFAS summary table


```{r summary_tables_remove_na}

pfas_data %>%
  select(seqn, pfdea, pfhxs, me_pfosa_acoh,	pfna, pfua,	pfdoa, n_pfoa, sb_pfoa, n_pfos,	sm_pfos) %>% 
  gather(key = "analyte", value = "value", pfdea:sm_pfos) %>% 
  group_by(analyte) %>% 
  na.omit() %>% 
  summarise(n = n(), 
            mean = mean(value),
            sd = sd(value)) %>%
  knitr::kable(digits = 2)


# Weighted

pfas_inputs <- as.data.frame(matrix(c(
  # CYCLE        COLUMN    COMMENT     WEIGHTS
   "2015-2016", "LBXPFDE", "LBXPFDEL", "WTSB2YR",
   "2015-2016", "LBXPFHS", "LBXPFHSL", "WTSB2YR",
   "2015-2016", "LBXMPAH", "LBXMPAHL", "WTSB2YR",
   "2015-2016", "LBXPFNA", "LBXPFNAL", "WTSB2YR",
   "2015-2016", "LBXPFUA", "LBXPFUAL", "WTSB2YR",
   "2015-2016", "LBXPFDO", "LBXPFDOL", "WTSB2YR",
   "2015-2016", "LBXNFOA", "LBXNFOAL", "WTSB2YR",
   "2015-2016", "LBXBFOA", "LBXBFOAL", "WTSB2YR",
   "2015-2016", "LBXNFOS", "LBXNFOSL", "WTSB2YR",
   "2015-2016", "LBXMFOS", "LBXMFOSL", "WTSB2YR"
), ncol = 4, byrow = TRUE), stringsAsFactors = FALSE)

names(pfas_inputs) <- c("cycle", "column", "comment_column", "weights_column")

nhanes_quantile(pfas_survey, pfas_inputs, quantiles = c(0.5, 0.95, 0.99))
nhanes_survey(svymean, pfas_survey, pfas_inputs, na.rm = TRUE)

```

### PFAS Boxplot

```{r boxplot_untransformed, echo = FALSE, eval = FALSE}
pfas_data %>%
  select(seqn, pfdea, pfhxs, me_pfosa_acoh,	pfna, pfua,	pfdoa, n_pfoa, sb_pfoa, n_pfos,	sm_pfos) %>% 
  gather(key = "analyte", value = "value", pfdea:sm_pfos) %>% 
  group_by(analyte) %>% 
  na.omit() %>% 
  ggplot(aes(x = analyte, y = value)) +
    geom_boxplot() 

```

```{r boxplot_log_scale}

pfas_data %>%
  select(seqn, pfdea, pfhxs, me_pfosa_acoh,	pfna, pfua,	pfdoa, n_pfoa, sb_pfoa, n_pfos,	sm_pfos) %>% 
  gather(key = "analyte", value = "value",  pfdea:sm_pfos) %>% 
  group_by(analyte) %>% 
  na.omit() %>% 
  ggplot(aes(x = analyte, y = value)) +
    geom_boxplot() +
     scale_y_log10()
```







### PFAS Ridge plot
```{r ridge_plot}

pfas_data %>%
  select(seqn, pfdea, pfhxs, me_pfosa_acoh,	pfna, pfua,	pfdoa, n_pfoa, sb_pfoa, n_pfos,	sm_pfos) %>% 
  gather(key = "analyte", value = "value", pfdea:sm_pfos) %>% 
  group_by(analyte) %>% 
  na.omit() %>% 
  filter(value < 10) %>% 
  ggplot(aes(x = value, y = analyte)) +
    ggridges::geom_density_ridges(scale = .85) +
    scale_x_log10()
    


    

```

### PFOS vs. PFOA scatterplot
```{r plot_pfoa_pfas, message = FALSE}
pfas_data %>% 
  ggplot(aes(x = n_pfoa, y = n_pfos)) + 
  geom_point(aes(color = riagendr), alpha = .5) +
  geom_smooth(se = TRUE) +
    scale_x_log10() +
    scale_y_log10()
```

### PFOS vs age acatterplot
```{r plot pfos_age, message = FALSE}
pfas_data %>% 
  ggplot(aes(x = ridageyr, y = n_pfos)) + 
  geom_point(aes(color = riagendr), alpha = .5) +
  geom_smooth(se = TRUE) +
    scale_y_log10()

```

## Body Mass

NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BMX_I.htm

### Load and Inspect BMI data

```{r load body mass data, message = FALSE}
bodymass_data <- nhanes_load_data("BMX_I", "2015-2016", demographics = TRUE)
as_tibble(bodymass_data)
```

### Inspect body mass data from 2015-2016
```{r examine_bmi, warning = FALSE, message = FALSE}
# bodymass_data %>%  nhanes_detection_frequency("BMXBMI", "BMXBMI", "WTMEC2YR") # not completely sure on weight
bodymass_data %>% nhanes_sample_size("BMXBMI", "BMXBMI", "WTMEC2YR")
bodymass_data %>%  nhanes_quantile("BMXBMI","BMXBMI", "WTMEC2YR", quantiles = c(0.5, 0.95))
```

### BMI Histogram, log transformed
```{r bmi_hist}
bodymass_data %>%  nhanes_hist("BMXBMI", "BMXBMI", "WTMEC2YR")
bodymass_data %>%  nhanes_hist("BMXBMI", "BMXBMI", "WTMEC2YR", transform = "log")

```

### BMI Scatterplot
```{r plot bmi_age, message = FALSE, warning = FALSE}
bodymass_data %>% 
  ggplot(aes(x = RIDAGEYR, y = BMXBMI)) + 
  geom_point(aes(color = RIAGENDR), alpha = .5) +
  geom_smooth(se = TRUE)

```

