---
title: "aamehs_rnhanes"
author: "Stephen Lewandowski"
date: "February 5, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RNHANES Exploration for AAMEHS

https://cran.r-project.org/web/packages/RNHANES/vignettes/introduction.html
https://www.r-bloggers.com/nhanes-made-simple-with-rnhanes/
https://datascienceplus.com/weight-loss-in-the-u-s-an-analysis-of-nhanes-data-with-tidyverse/


Install RNHANES package (first time - need development version for most recent cycles)
```{r install_package}
# library(devtools)
# install_github("silentspringinstitute/RNHANES")
```

```{r packages, message = FALSE}
library(tidyverse)
library(RNHANES)
```


NHANES files and the comprehensive variable list
```{r load_file_names, cache = TRUE, message = FALSE}
files <- nhanes_data_files()  #1,328 obs of 10 variables
variables <- nhanes_variables()  #slow, takes ~5 minutes (48,398 obs of 10 variables)

```

```{r explore_files}
files %>% count(cycle) %>% View
files %>% count(component)
```

## Search files and variables
```{r search_files_options}
nhanes_search(files, "environmental phenols")
nhanes_search(files, "pesticides", component == "laboratory", cycle == "2003-2004") 
nhanes_search(files, "", cycle == "2003-2004") %>% head
nhanes_search(variables, "triclosan") %>% head
nhanes_search(variables, "DDT", data_file_name == "LAB28POC") %>% head
nhanes_search(variables, "", data_file_name == "EPH_E") %>% head
```


## Download data files

specify the name and cycle year, 
Saves the files to a temporary directory by default. You can optionally set where you want the files to be downloaded.
```{r load, eval = FALSE}
nhanes_load_data("EPH", "2007-2008", cache = "./nhanes_data", demographics = TRUE, recode = TRUE)
```


Example of data loaded from multiple files/cycle years. Download all files that contain a "PFOS" variable.
```{r search_pfos, eval = FALSE}
?RNHANES::nhanes_search
results <- nhanes_search(variables, "PFOS")
results$cycle
pfos <- nhanes_load_data(results$data_file_name, results$cycle, demographics = TRUE, recode = TRUE) #slow; large list, 9 elements, 12 MB

pfos
```

## Load and Inspect PFAS data
NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/PFAS_H.htm
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/PFAS_I.htm

```{r load_pfas}
pfas_data <- nhanes_load_data("PFAS_I", "2015-2016", demographics = TRUE)
as_tibble(pfas_data)
```

Inspect PFAS data from 2015-2016
```{r examine_pfas}
pfas_data %>%  nhanes_detection_frequency("LBXNFOA", "LBDNFOAL")
pfas_data %>% nhanes_sample_size("LBXNFOA","LBDNFOAL")
pfas_data %>%  nhanes_quantile("LBXNFOA", "LBDNFOAL", quantiles = c(0.5, 0.95))
```

Histogram, log transformed
```{r pfas_hist}
pfas_data %>%  nhanes_hist("LBXNFOA", "LBDNFOAL")
pfas_data %>%  nhanes_hist("LBXNFOA", "LBDNFOAL", transform = "log")

```


```{r plot_pfoa_pfas, message = FALSE}
pfas_data %>% 
  ggplot(aes(x = LBXNFOA, y = LBXNFOS)) + 
  geom_point(aes(color = RIAGENDR), alpha = .5) +
  geom_smooth(se = TRUE)
```

```{r plot pfos_age, message = FALSE}
pfas_data %>% 
  ggplot(aes(x = RIDAGEYR, y = LBXNFOS)) + 
  geom_point(aes(color = RIAGENDR), alpha = .5) +
  geom_smooth(se = TRUE)

```


## Load and Inspect BMI data

NHANES Codebook References:
https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/BMX_I.htm

```{r load body mass data}
bodymass_data <- nhanes_load_data("BMX_I", "2015-2016", demographics = TRUE)
as_tibble(bodymass_data)
```

Inspect body mass data from 2015-2016
```{r examine_bmi}
bodymass_data %>%  nhanes_detection_frequency("BMXBMI", "BMXBMI", "WTMEC2YR") # not completely sure on weight
bodymass_data %>% nhanes_sample_size("BMXBMI", "BMXBMI", "WTMEC2YR")
bodymass_data %>%  nhanes_quantile("BMXBMI","BMXBMI", "WTMEC2YR", quantiles = c(0.5, 0.95))
```

Histogram, log transformed
```{r bmi_hist}
bodymass_data %>%  nhanes_hist("BMXBMI", "BMXBMI", "WTMEC2YR")
bodymass_data %>%  nhanes_hist("BMXBMI", "BMXBMI", "WTMEC2YR", transform = "log")

```

```{r plot bmi_age, message = FALSE}
bodymass_data %>% 
  ggplot(aes(x = RIDAGEYR, y = BMXBMI)) + 
  geom_point(aes(color = RIAGENDR), alpha = .5) +
  geom_smooth(se = TRUE)

```
